## 1. Load Data

Load data from datasets.zip file.

## 2. EDA - Data Cleaning and Initial Analysis

### 2.1 Summarize Resource Dataset

```{r income-EDA-start, include=T , eval=T}
# Select and rename data from dataset as income_0, which contains income, children
df_income <- income_dateset %>% 
  select(Total.Household.Income,
         Members.with.age.5...17.years.old
         ) %>% 
  rename(income = Total.Household.Income, 
         children = Members.with.age.5...17.years.old
         )

# Summarize df_income
xskim   <- skim(df_income)
print(xskim)

```

### 2.2 Dataset Preparation

**Dataset Descriptions:**

-   **income_0**: Initial dataset (`df_income`).

-   **income_1**: Dataset after removing custom outliers.

-   **income_2**: Log-transformed income with children as a feature.

```{r Prepare 3 dataset, include=T , eval=T}
# Drop missing income  
income_0 <- df_income |> drop_na(income) # income_0 is data with outliers

# Outliers Removal using IQR method (5xIQR, upper limit 2000000)
# Call `remove_outliers_IQR`, using a multiplier of 5 and setting an upper cap of 2,000,000.
income_1 <- remove_outliers_IQR(df=income_0,
                                col="income",
                                multiplier = 5,
                                upper_cap = 2000000) 

# log(income) prepare
income_2 <- data.frame(
  income = log(income_0$income),  
  children = income_0$children          
  )

# Generate and print summary statistics for the datasets: 
# income_0 (initial dataset), 
# income_1 (dataset without custom outliers), 
# and income_2 (log-transformed income with children).

xskim0 <- skim(income_0)
xskim1 <- skim(income_1)
xskim2 <- skim(income_2)

print(xskim0)
print(xskim1)
print(xskim2)

```

### 2.3 Visualisation Of Three Datasets

```{r Visualisation Of Three Datasets, include=T, eval=T, echo=F}
# Identify removed outliers
removed_rows <- setdiff(income_0, income_1)  # Gets the removed outlier

# Calculate the number and proportion of outliers
num_outliers <- nrow(removed_rows)
total_rows <- nrow(income_0)
outlier_percentage <- (num_outliers / total_rows) * 100  # Calculated percentage
outlier_text <- paste0("Marked ", 
                       num_outliers,
                       " outliers (",
                       round(outlier_percentage, 2), "%)"
                       )

# Plot a scatter plot, marking normal values (black) and outliers (red)
plot(y=income_0$income / 1000000, x=income_0$children,
     xlab = "Children", 
     ylab = "Household Income (in 1 Million)", 
     main = "Household Income Distribution Based on Number of Children",
     pch = 4, col = "black")  # The black X represents normal data

# Mark outliers in red
points(y=removed_rows$income / 1000000, x=removed_rows$children,
       pch = 4, col = "red")  # The red X represents outliers

# Text annotation of the number and proportion of outliers
text(y = max(income_0$income / 1000000) * 0.8, 
     x = max(income_0$children) * 0.6, 
     labels = outlier_text, 
     col = "red", 
     font = 2, 
     cex = 1)  # Font set as bold and size fit

# Add a legend to the plot
legend("topright", legend = c("income_1", "outliers"),
       col = c("black", "red"), pch = 4, bty = "n", cex = 0.8)

```

The figure indicates that the outliers seem far more numerous than they truly are. How we define outliers could also reshape our dataset thus leading to different outcomes.

```{r visualization-EDA-log(income), include=T,eval=T,echo=F}
# plot scatter
plot(y=income_2$income, x=income_2$children,
     xlab = "Children", 
     ylab = "Log(Income)", 
     main = "Log(Income) Distribution Based on Number of Children",
     pch = 4, col = "black", 
     xlim = c(0, max(income_2$children) + 1))  # Set the X-axis range

# define axis length
axis(1, at=0:max(income_2$children), labels=0:max(income_2$children))
```

Taking the logarithm of income can effectively reduce the impact of outliers and linearise relationships between children and income while also giving it more chance to be misunderstood by unprofessional.

\newpage

## 3. Linear Regression

### 3.1 Training Dataset And Testing Dataset Set-Up Function

This split_data function splits a given dataset into training and test sets based on a specified training ratio, ensuring that the training set contains a random subset of the data.

```{r training-data-and-testing-data-setup, include=T, eval=T}
# Set a seed for reproducibility and to minimize RAM usage
set.seed(62380486) 
# Function to split data into training and test sets
split_data <- function(data, train_ratio = 0.8) {
  # validate train_ratio range
  if (train_ratio <= 0 || train_ratio >= 1) {
  stop("Error: train_ratio must be between 0 and 1 (exclusive).")
}
  # Randomly select the specified percentage of indices for the training set
  train_ind <- sample(1:nrow(data), 
                      size = floor(train_ratio * nrow(data)),
                      replace = FALSE)
  
  # Use the remaining indices for the test set
  test_ind <- setdiff(1:nrow(data), train_ind)
  
  # Create training data using the selected indices
  train_data <- data[train_ind, , drop = FALSE]
  rownames(train_data) <- NULL

  # Create test data using the remaining indices
  test_data <- data[test_ind, , drop = FALSE]
  rownames(test_data) <- NULL
  
  # Return both training and test data as a list
  return(list(train = train_data, test = test_data))
}


```

### 3.2 Model Training Function Set-Up

```{r Train-a-Model-with-the-parameterized-dataset,include=T, eval=T}
# Function to split data and fit linear regression
fit_model <- function(data, response, predictor, train_ratio = 0.8) {
  # Split data by function split_data
  xsplit_data <- split_data(data, train_ratio)
  xtrain_data <- xsplit_data$train
  xtest_data <- xsplit_data$test
  
  # Dynamically create formula: response ~ predictor
  formula <- as.formula(paste(response, "~", predictor),)
  
  # Fit the linear model
  model <- lm(formula, data = xtrain_data)
 
  # Return model and data
  return(list(model = model, train = xtrain_data, test = xtest_data))
}


```

The return statement in this function outputs a list containing three components:

-   the fitted linear model ( model ),

-   the training data ( train ),

-   and the test data ( test ).

This structure emphasizes that the function provides not only the model but also the datasets used for training and testing, which may help prevent misunderstandings in subsequent calls where users might expect a direct output of the model alone.

### 3.3 Model Training

We will train three models to evaluate and compare their performance.

`xmodel_0` will be trained using income_0 .

`xmodel_1` will be trained using income_1 .

`xmodel_2` will be trained using income_2 .

```{r train-models-with-outliers-and-without-outliers, include=T,eval=T,echo=T}
xmodel_0 <- fit_model(data=income_0, 
                      response = "income", 
                      predictor = "children", 
                      train_ratio = 0.8)

xmodel_1 <- fit_model(data=income_1,
                      response = "income",
                      predictor = "children",
                      train_ratio = 0.8)

xmodel_2 <- fit_model(data=income_2,
                      response = "income",
                      predictor = "children",
                      train_ratio = 0.8)

```

### 3.4 Model Comparison

In this section, we define a `compare_models` function to enhance the efficiency of comparing the three models mentioned above.

```{r Model-Comparison-function, include=TRUE, eval=TRUE}
compare_models <- function(..., model_names = NULL) {
  # magic parameter ... allow this func to accept unfixed numerous models
  models <- list(...)  
  
  num_models <- length(models)
  
  # Default model names if not provided
  if (is.null(model_names)) {
    model_names <- paste0("Model_", seq_len(num_models))
  }
  
  # Initialize results list
  results <- list()
  
  for (i in seq_len(num_models)) {
    model_summary <- summary(models[[i]])
    
    # Extract metrics
    intercept <- sprintf("%.2f", round(coef(model_summary)[1, 1], 2))
    slope <- sprintf("%.2f", round(coef(model_summary)[2, 1], 2))
    r_squared <- formatC(model_summary$r.squared,
                         format = "e",
                         digits = 4
                        )
    adj_r_squared <- formatC(model_summary$adj.r.squared,
                             format = "e",
                             digits = 4
                            )
    residual_se <- sprintf("%.2f", round(model_summary$sigma, 2))
    f_stat <- sprintf("%.4f", round(model_summary$fstatistic[1], 4))
    p_value <- formatC(pf(model_summary$fstatistic[1],
                                           model_summary$fstatistic[2],
                                           model_summary$fstatistic[3],
                                           lower.tail = FALSE
                                          ),
                                       format = "e",
                                       digits = 6
                       )
    
  
    # Store in list
    results[[model_names[i]]] <- c(Intercept = intercept,
                                   Slope = slope,
                                   R_squared = r_squared,
                                   Adj_R_squared = adj_r_squared,
                                   Residual_SE = residual_se,
                                   F_statistic = f_stat,
                                   p_value = p_value)
  }
  
  # Combine into data frame
  comparison_df <- as.data.frame(do.call(cbind, results))
  comparison_df <- cbind(Metric = rownames(comparison_df), comparison_df)
  rownames(comparison_df) <- NULL
  
  return(comparison_df)
}

# Compare models 
comparison_table <- compare_models(xmodel_0$model,
                                   xmodel_1$model,
                                   xmodel_2$model,
                                   model_names = c("xmodel_0",
                                                   "xmodel_1",
                                                   "xmodel_2")
                                   )
print(comparison_table)
```

### 3.3 Model Visualization

```{r unified-visualization-function, include=T,eval=T,echo=F}
# Define a unified visualization function
plot_model_fit <- function(model, data, response, predictor) {
  # Extract the predicted values from the model
  data$predicted <- predict(model, newdata = data)
  
  # Create the visualization plot
  ggplot(data, aes(x = .data[[predictor]], y = .data[[response]])) +
    geom_point(alpha = 0.5) +  # Plot scatter points
    geom_line(aes(y = predicted), color = "blue", linewidth = 1) +  # Plot the fitted line
    scale_y_continuous(
      expand = expansion(mult = c(0.05, 0.05)),
      labels = function(x) {
        ifelse(x > 2000000, paste0(x / 1e6, "M"),  # Convert values greater than 2 million to millions
               ifelse(x > 10000, paste0(x / 1000,"K"), x))  # Convert values greater than 10,000 to thousands
      }
    ) +
    labs(
      title = paste("Model Fit Line in Model:", deparse(substitute(model))), 
      x = predictor, 
      y = ifelse(max(data[[response]]) > 2000000, 
                  paste0(response, " (unit: millions)"),  # Label in millions
                  ifelse(max(data[[response]]) > 10000, 
                         paste0(response, " (unit: thousands)"), response))  # Label in thousands
    ) +
    theme_minimal()
}

# Use the function to visualize the fit line for each model
plot_model_fit(xmodel_0$model, income_0, "income", "children")
plot_model_fit(xmodel_1$model, income_1, "income", "children")
plot_model_fit(xmodel_2$model, income_2, "income", "children")



```

\newpage

## 4. Model Explanation

### 4.1 Model Form

The specific form of this model is expressed as: $income = b_0 + b_1 \cdot children + \epsilon$, and could be given as: $income \approx b_0 + b_1 \cdot children$ when $\epsilon$ is acceptable in practice.

-   $b_0$ (intercept): Represents the predicted income when the number of children is zero.

-   $b_1$ (slope): Indicates the change in income for each additional child.

-   $\epsilon$ (a mean-zero random error term): The error term is a catch-all for what we miss with this simple model: the true relationship is probably not linear, there may be other variables that cause variation in $income$, and there may be measurement error. We typically assume that the error term is independent of $children$, and $\epsilon \sim \mathcal{N}(0,\sigma^2)$. (James et al., 2021, p. 63)

-   Each computed on the basis of a separate random set of observations, is different. (James et al., 2021, p. 64) Using a different seed number, specifically set to `62380486` at the outset of the dataset splits, the coefficients $b_0$ and $b_1$ exhibit variability within certain intervals.

For clear, we present the estimated coefficients $\hat{b}_0$ and $\hat{b}_1$ from `xmodel_0` ,`xmodel_1`and `xmodel_2` .

```{r print-b0-and-b1, include=T,eval=T,echo=F}
b0_0 <- coef(xmodel_0$model)[1]  
b1_0 <- coef(xmodel_0$model)[2] 
cat(paste0("Model name : xmodel_0 \n b0 = ", b0_0, ", b1 = ", b1_0, "\n"))

b0_1 <- coef(xmodel_1$model)[1]  
b1_1 <- coef(xmodel_1$model)[2] 
cat(paste0("Model name : xmodel_1 \n b0 = ", b0_1, ", b1 = ", b1_1, "\n"))

b0_2 <- coef(xmodel_2$model)[1]  
b1_2 <- coef(xmodel_2$model)[2] 
cat(paste0("Model name : xmodel_2 \n b0 = ", b0_2, ", b1 = ", b1_2, "\n"))


```

### 4.2 Prediction Mapping

`pred_income` Function Description:

The `pred_income` function calculates predicted income based on a linear model. It takes three arguments:

-   `b0`: The intercept of the model.
-   `b1`: The slope of the model, indicating the change in income for each unit change in `n`.
-   `n`: A numeric vector representing the independent variable (e.g., number of children).

The function initializes a numeric vector to store the results and iterates over the elements of `n`, applying the linear equation $\text{income} = b_0 + b_1 \times n[i]$ to compute the predicted income for each value in `n`. Finally, it returns the vector of predicted income values.

```{r pred_income function, include=TRUE,eval=TRUE}
pred_income <- function(b0, b1, n) {
  
  result <- numeric(length(n))  # Initialize a numeric vector to store results
  
  for (i in seq_along(n)) {     # Iterate over the indices of n
    result[i] <- b0 + b1 * n[i]  # Calculate income and store it
  }
  return(result)                 # Return the result
}

```

```{r xtrain_pred, include=TRUE,eval=T}
# Define n as a vector
n <- 0:8  # Number of children (from 0 to 8)
alpha <- 0.1
# Calculate predicted income for each n and round to two decimal places
pred_0 <- round(pred_income(b0_0, b1_0, n), 2)
pred_1 <- round(pred_income(b0_1, b1_1, n), 2)
pred_2 <- round(pred_income(b0_2, b1_1, n), 2)

```

### 4.3 interval for $\hat{b}_1$

To calculate the confidence intervals for $\hat{b}_1$, we apply the formula provided in Week 3 of the lecture notes (Li, 2025). The formula allows us to estimate the range within which the true value of $b_1$ is likely to fall.

$$
b_1 \in [ \hat{b}_1-t_{1-\frac{\alpha}{2}}(n-2) \cdot se(\hat{b}_1),\hat{b}_1+t_{1-\frac{\alpha}{2}}(n-2) \cdot se(\hat{b}_1)]
$$

where

$t_r(k)$ is the r-quantile of a t-distribution with $k$ degrees of freedom and

$se(\hat{b}_1)= \sqrt{\frac{1}{n-2} \cdot\frac{\sum_{i=1}^{n}(y_i-\hat{y}_i)^2}{\sum_{i=1}^{n}(x_i-mean(\underline{x}))^2}}$

Thus we define a function to calculate the confidence with probability $100 \cdot (1-\alpha) \%$.

The `confidence_interval_b1` function computes the confidence interval for the estimated slope coefficient $b_1$ of a linear regression model.

Parameters:

-   `y`: A numeric vector representing the dependent variable, or response.
-   `x`: A numeric vector representing the independent variable, or predictor.
-   `b1_hat`: The estimated slope coefficient from given linear regression model.
-   `alpha`: The significance level (default is 0.05), which determines the confidence level as $100 \cdot (1 - \alpha) \%$.

```{r confidence_interval_b1-calculation-function set-up, include=TRUE,eval=TRUE}
# Compute the confidence interval for b1
confidence_interval_b1 <- function(y, x, b1_hat, alpha=0.05) {
  n <- length(y)
  
  # predict value y_hat
  y_hat <- predict(lm(y ~ x))
  #  se(b1_hat)
  residuals <- y - y_hat
  se_b1_hat <- sqrt(sum(residuals^2) / (n - 2)) / sqrt(sum((x - mean(x))^2))
  
  #  r-quantile of a t-distribution with n-2 degrees of freedom
  t_value <- qt(1 - alpha/2, df = n - 2)
  
  # interval for b1_hat
  lower_bound <- b1_hat - t_value * se_b1_hat
  upper_bound <- b1_hat + t_value * se_b1_hat
  
  return(c(lower_bound, upper_bound))
}

```

```{r Calculate-Prediction-Intervals-for-b1_hat-of-3-Models, include=TRUE,eval=T,echo=T}
xintervals_b1 <- data.frame(
  Model = c("xmodel_0", "xmodel_1", "xmodel_2"),
  Lower_Bound = c(
    confidence_interval_b1(y=xmodel_0$train$income,
                           x=xmodel_0$train$children,
                           b1_hat=coef(xmodel_0$model)[2],
                           alpha = alpha
                           )[1],
    
    confidence_interval_b1(y=xmodel_1$train$income,
                           x=xmodel_1$train$children,
                           b1_hat=coef(xmodel_1$model)[2],
                           alpha = alpha
                           )[1],
    
    confidence_interval_b1(y=xmodel_2$train$income,
                           x=xmodel_2$train$children,
                           b1_hat=coef(xmodel_2$model)[2],
                           alpha = alpha
                           )[1]
  ),
  
  Upper_Bound = c(
    confidence_interval_b1(y=xmodel_0$train$income, 
                           x=xmodel_0$train$children, 
                           b1_hat=coef(xmodel_0$model)[2],
                           alpha = alpha
                           )[2],
    
    confidence_interval_b1(y=xmodel_1$train$income, 
                           x=xmodel_1$train$children, 
                           b1_hat=coef(xmodel_1$model)[2], 
                           alpha = alpha
                           )[2],
    
    
    confidence_interval_b1(y=xmodel_2$train$income,
                           x=xmodel_2$train$children, 
                           b1_hat=coef(xmodel_2$model)[2],
                           alpha = alpha
                           )[2]
  )
)

# knitr::kable() show result in table
knitr::kable(xintervals_b1, caption = "Prediction Intervals for b1 of 3 Models")

```

Testing $ð‘_1$ for significance

"If 0 is contained in the confidence interval, then we say that $b_1$ is significant at level $\alpha$. (lecture notes page 44, Li, T. (2025).

Consequently, the coefficients for both `xmodel_0` and `xmodel_1` are statistically significant at the $\alpha=0.1$ level, whereas the $b_1$ coefficient for `xmodel_2` is statistically insignificant.

#### Summarises in tables

```{r Display-each-results-table,include=T,eval=T,echo=F}
# Create a results data frame for each model
results_0 <- data.frame(
  n = n,
  Pred_Income = pred_0,
  b1_hat = coef(xmodel_0$model)[2],
  Lower_Bound = confidence_interval_b1(y=xmodel_0$train$income, x=xmodel_0$train$children, b1_hat=coef(xmodel_0$model)[2], alpha = alpha)[1],
  Upper_Bound = confidence_interval_b1(y=xmodel_0$train$income, x=xmodel_0$train$children, b1_hat=coef(xmodel_0$model)[2], alpha = alpha)[2],
  row.names = NULL  # Set row names to NULL to avoid warnings
)

results_1 <- data.frame(
  n = n,
  Pred_Income = pred_1,
  b1_hat = coef(xmodel_1$model)[2],
  Lower_Bound = confidence_interval_b1(y=xmodel_1$train$income, x=xmodel_1$train$children, b1_hat=coef(xmodel_1$model)[2], alpha = alpha)[1],
  Upper_Bound = confidence_interval_b1(y=xmodel_1$train$income, x=xmodel_1$train$children, b1_hat=coef(xmodel_1$model)[2], alpha = alpha)[2],
  row.names = NULL  # Set row names to NULL to avoid warnings
)

results_2 <- data.frame(
  n = n,
  Pred_Income = pred_2,
  b1_hat = coef(xmodel_2$model)[2],
  Lower_Bound = confidence_interval_b1(y=xmodel_2$train$income, x=xmodel_2$train$children, b1_hat=coef(xmodel_2$model)[2], alpha = alpha)[1],
  Upper_Bound = confidence_interval_b1(y=xmodel_2$train$income, x=xmodel_2$train$children, b1_hat=coef(xmodel_2$model)[2], alpha = alpha)[2],
  row.names = NULL  # Set row names to NULL to avoid warnings
)



```

```{r knitr-Display-each-results-table,include=T,eval=T,echo=F}
knitr::kable(results_0, caption = "Prediction Intervals and Predicted Income for xmodel_0")
knitr::kable(results_1, caption = "Prediction Intervals and Predicted Income for xmodel_1")
knitr::kable(results_2, caption = "Prediction Intervals and Predicted Income for xmodel_2")
```

The confidence interval will not change as `n` changes, because the $\hat{b}_1$ relies on its model's train dataset, rather than any simple point.

### 4.4 Prediction Intervals

A $100 \cdot (1-\alpha)\%$ prediction interval for $x^*$ is $[\hat{y^*-\tau},\hat{y^*-\tau}]$,where

$$
\tau = t_{1-\frac{\alpha}{2}} \cdot(n-2) \cdot \sqrt{\frac{RSS}{n-2}} \cdot \sqrt{1+\frac{1}{n}+\frac{(x^*-mean(\underline{x}))^2}{\sum_{i=1}^{n}(x_i-mean(\underline{x}))^2}}
$$

Thus we can define a function to compute the prediction interval for a given $x^*$

```{r prediction_interval-function, include=TRUE,eval=TRUE,echo=TRUE}
# Function to compute prediction interval using a pre-trained model
# 
# This function calculates prediction intervals for new data points based on
# a fitted linear regression model
# Parameters:
#   model: A fitted linear regression model (lm object)
#   x_train: Vector of predictor values from the training set
#   x_star: Vector of predictor values for which to compute prediction intervals
#   alpha: Significance level (default = 0.1 for 90% prediction intervals)
# Returns:
#   A data frame containing the predictor values, predicted responses,\n and lower/upper bounds
prediction_interval_model <- function(model, x_train, x_star, alpha = 0.1) {
  # Get the sample size of training data
  n <- length(x_train)
  
  # Extract residual standard error from the model
  residuals <- model$residuals
  RSS <- sum(residuals^2)  # Residual sum of squares
  sigma_hat <- sqrt(RSS / (n - 2))  # Estimate of error standard deviation
  
  # Compute mean and sum of squared deviations for training predictors
  x_bar <- mean(x_train)
  Sxx <- sum((x_train - x_bar)^2)  # Sum of squared deviations
  
  # Calculate critical t-value for the given confidence level
  t_value <- qt(1 - alpha / 2, df = n - 2)
  
  # Extract predictor variable name from the model formula to create proper newdata
  formula_terms <- terms(model)
  predictor_name <- attr(formula_terms, "term.labels")[1]
  newdata <- data.frame(x_star)
  names(newdata) <- predictor_name
  
  # Generate point predictions for new data points
  y_star_hat <- predict(model, newdata = newdata)
  
  # Calculate prediction interval width (tau) for each new data point
  # This accounts for three sources of uncertainty:
  # 1. Inherent variability (1)
  # 2. Uncertainty in estimating the mean (1/n)
  # 3. Distance of x_star from the mean of training data ((x_star - x_bar)^2/Sxx)
  tau <- t_value * sigma_hat * sqrt(1 + (1 / n) + ((x_star - x_bar)^2) / Sxx)
  
  # Calculate lower and upper bounds of prediction intervals
  lower_bound <- y_star_hat - tau
  upper_bound <- y_star_hat + tau
  
  # Return results as a data frame
  return(data.frame(x_star = x_star,
                    y_hat = y_star_hat,
                    lower = lower_bound,
                    upper = upper_bound))
}

# Function to evaluate prediction interval coverage on test data
#
# This function checks what percentage of actual test values fall within the calculated prediction intervals
# Parameters:
#   model: A fitted linear regression model (lm object)
#   x_train: Vector of predictor values from the training set
#   x_test: Vector of predictor values from the test set
#   y_test: Vector of response values from the test set
#   alpha: Significance level (default = 0.1 for 90% prediction intervals)
#   model_name: Name identifier for the model (useful when comparing multiple models)
# Returns:
#   A list containing the coverage percentage and detailed results for each test point
check_prediction_coverage_model <- function(model, x_train, x_test, y_test, alpha = 0.10, model_name = "Model") {
  # Validate inputs - check for missing values in test data
  if (any(is.na(x_test)) || any(is.na(y_test))) {
    stop("Error: x_test or y_test contains NA values.")
  }
  
  # Calculate prediction intervals for all test points
  intervals <- prediction_interval_model(model, x_train, x_test, alpha)
  
  # Determine which test points fall within their prediction intervals
  within_interval <- (y_test >= intervals$lower) & (y_test <= intervals$upper)
  
  # Calculate the percentage of test points within intervals
  percent_within <- mean(within_interval, na.rm = TRUE) * 100
  
  # Print a summary of the coverage results with model identification
  cat(sprintf("%s: Percentage of test points within %.0f%% prediction interval: %.2f%%\n", 
              model_name, (1 - alpha) * 100, percent_within))
  
  # Return detailed results
  return(list(percent_within = percent_within,
              details = data.frame(x_test = x_test,
                                   y_test = y_test,
                                   lower = intervals$lower,
                                   upper = intervals$upper,
                                   within = within_interval)))
}


# Example usage: Evaluate prediction interval coverage for model 0
# This model predicts income based on number of children
coverage_result0 <- check_prediction_coverage_model(
  model = xmodel_0$model,  # The fitted linear model
  x_train = xmodel_0$train$children,  # Predictor values from training set
  x_test = xmodel_0$test$children,  # Predictor values from test set
  y_test = xmodel_0$test$income,  # Actual response values from test set
  alpha = 0.1,  # For 90% prediction intervals
  model_name = "xmodel_0"  # Model identifier
)


# Evaluate prediction interval coverage for model 1
coverage_result1 <- check_prediction_coverage_model(
  model = xmodel_1$model, 
  x_train = xmodel_1$train$children,
  x_test = xmodel_1$test$children, 
  y_test = xmodel_1$test$income,
  alpha = 0.1,
  model_name = "xmodel_1"
)

# Evaluate prediction interval coverage for model 2
# This model uses a log transformation of income
coverage_result2 <- check_prediction_coverage_model(
  model = xmodel_2$model, 
  x_train = xmodel_2$train$children,
  x_test = xmodel_2$test$children, 
  y_test = xmodel_2$test$income,
  alpha = 0.1,
  model_name = "xmodel_2"
)

```

```{r visualisation-function-coverage, include=TRUE,eval=TRUE,echo=F}
# Function to visualize prediction interval coverage
# 
# This function creates a plot showing test points and their prediction intervals
# Parameters:
#   coverage_result: The result from check_prediction_coverage_model function
#   title: Plot title (default is "Prediction Interval Coverage")
#   x_lab: Label for x-axis (default is "Predictor")
#   y_lab: Label for y-axis (default is "Response")
#   point_size: Size of the plotted points (default is 2)
#   alpha: Transparency of prediction interval bands (default is 0.3)
# Returns:
#   A ggplot object showing the prediction intervals and test points
plot_prediction_coverage <- function(coverage_result, 
                                    title = "Prediction Interval Coverage",
                                    x_lab = "Predictor", 
                                    y_lab = "Response",
                                    point_size = 2,
                                    alpha = 0.3) {
 
  # Extract details from coverage result
  details <- coverage_result$details
  
  # Create the plot
  p <- ggplot(details, aes(x = x_test)) +
    # Add prediction interval ribbon
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "lightblue", alpha = alpha) +
    # Add points, colored by whether they fall within the interval
    geom_point(aes(y = y_test, color = within), size = point_size) +
    # Customize colors
    scale_color_manual(values = c("FALSE" = "red", "TRUE" = "darkblue"),
                      name = "Within interval",
                      labels = c("FALSE" = "Outside", "TRUE" = "Inside")) +
    # Add labels and title
    labs(title = title,
         subtitle = paste0("Coverage: ", round(coverage_result$percent_within, 1), "%"),
         x = x_lab,
         y = y_lab) +
    # Add theme elements
    theme_minimal() +
    theme(legend.position = "bottom")
  
  return(p)
}

# Example usage for each model
# Plot prediction interval coverage for model 0
plot0 <- plot_prediction_coverage(
  coverage_result0,
  title = "Model 0: Prediction Interval Coverage",
  x_lab = "Number of Children",
  y_lab = "Income"
)

# Plot prediction interval coverage for model 1
plot1 <- plot_prediction_coverage(
  coverage_result1,
  title = "Model 1: Prediction Interval Coverage",
  x_lab = "Number of Children",
  y_lab = "Income"
)

# Plot prediction interval coverage for model 2
plot2 <- plot_prediction_coverage(
  coverage_result2,
  title = "Model 2: Prediction Interval Coverage (log transformation)",
  x_lab = "Number of Children",
  y_lab = "Income"
)

# Display the plots
print(plot0)
print(plot1)
print(plot2)



```
